---
title: "4/07"
author: Olga Sironi
output: html_document
---

## R Functions
object: everything is an object
functions: everything that is happening in R is the result of a function call.

```{r}
n <- 100
u <-runif(n)
n
```

```{r}
?runif
```

## Functions
Function principles:

The value of a function call is uniquely defined by the value of the argument.

The effect of a function call is simply the returned value.


```{r}
runif(10)
```
It's different in this case. They are random number generating functions.
```{r}
runif(10)
```

It's important to look at the anatomy of the function in Ch. 6.1:
name, arguments, body and returned value.

R returns automatically the last thing evaluated.

## Creating functions
Write a function that adds 2 numbers together

```{r}
add <- function(a,b){
  a + b
}
add (1,2)
```


```{r}
add <- function(a,b){
  if (a==1) return(b)
}
add (1,2)
```

```{r}
add <- function(a,b){
  if (a==1) return(b)
  a + b
}

addb <- function(a, b=1){
  a + b
}
addb(1)
```


```{r}
add <- function(a,b){
  if (a==1) return(b)
  a + b
}

addb <- function(a, b=1){
  a + b
}
addb(1,2)
```

# The lazy rule
Any timke you repeat something more than 2 times, consider the use of functions.

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10)
) 
df$a <- (df$a -min(df$a))/max(df$a - min(df$a))

df$a
```

```{r}
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10)
) 
df$a <- (df$a -min(df$a))/max(df$a - min(df$a))
df$b <- (df$b -min(df$b))/max(df$b - min(df$b))
df$c <- (df$c -min(df$c))/max(df$c - min(df$c))

df
```

```{r}
set.seed(123)
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10)
) 
df$a <- (df$a -min(df$a))/max(df$a - min(df$a))
df$b <- (df$b -min(df$b))/max(df$b - min(df$b))
df$c <- (df$c -min(df$c))/max(df$c - min(df$c))

df
rescale <- function(x){
  (x-min(x))/(max(x)-min(x))
  
}
```


```{r}
set.seed(123)
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10)
) 
df$a <- (df$a -min(df$a))/max(df$a - min(df$a))
df$b <- (df$b -min(df$b))/max(df$b - min(df$b))
df$c <- (df$c -min(df$c))/max(df$c - min(df$c))

rescale <- function(x){
  (x-min(x))/(max(x)-min(x))
  
}

df$a <- rescale(df$a)
df$b <- rescale(df$b)
df$c <- rescale(df$c)

apply(df,2,rescale)
```

```{r}
set.seed(123)
df <- data.frame(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10)
) 
df$a <- (df$a -min(df$a))/max(df$a - min(df$a))
df$b <- (df$b -min(df$b))/max(df$b - min(df$b))
df$c <- (df$c -min(df$c))/max(df$c - min(df$c))

rescale <- function(x){
  (x-min(x))/(max(x)-min(x))
  
}

df$a <- rescale(df$a)
df$b <- rescale(df$b)
df$c <- rescale(df$c)

df_rescaled <- apply(df,2, rescale)
df_rescaled
```

## Environments

It controls how R finds the value associated with a name.

R creates a call environment and within the arguments it does something in the value. So for instanve we have a=2 and B=2 and a+b=4. Around this environment, there's the global environment. If R tries to add a+b and doesn't have a value for them in the call environment, it will have a look in the global environment for a value associated in the name, without returning any error. If it cannot find a value here, it finds them in the parent environment.

```{r}
add <- function(a){
  a + h
}
h <- 2
add(3)

```

## Generic functions

R is an object-oriented programming language, but not like Java or C++, that have an encapsulated OOP.
It has functional-oriented object language.
This means that a method has different behaviors depending on the class of the object.

```{r}
letters  <- factor (sample(letters[1:4], 10, replace = T))
numbers <- rnorm(10)
letters <- base::letters
letters
```

```{r}
let  <- factor (sample(letters[1:4], 10, replace = T))
num <- rnorm(10)
let
```
```{r}
num 
```

The function behaves differently depending on the class of the objects.

```{r}
class(num )
```

For generic functions, R looks at the first argument.

```{r}
let  <- factor (sample(letters[1:4], 10, replace = T))
num <- rnorm(10)
class(let)
print.default(let)

as.numeric(let)
```

```{r}
plot(let)
```
```{r}
plot(num)
```

It depends on the class of the elements.

## Exercise
Create a function that plots an image of a matrix

```{r}
a <- matrix(rnorm(100), 10, 10)
plot(a)
image(a)
```

To have this result by writing plot I add class and define function for it.

```{r}
a <- matrix(rnorm(100), 10, 10)
plot(a)
image(a)

plot.matrix <- function(x){
  image(x)
}
plot(a)
```
